This thesis is the result of my graduation project and concludes the part-time
Master programme Software Engineering at the University of Amsterdam.

The project has been performed at VNU Vacature Media, the company at which I am
employed at. VNU Vacature Media is a division of De Persgroep Nederland since
May 2012. The company operates in the on-line recruitment market. It owns the
largest job boards in The Netherlands and Belgium: Intermediair,
Nationale Vacaturebank, Vacature.com, and References.be. The software
supporting these products are mostly open-source projects, that is why the
survivability of open-source software projects came to my interest in the choice
of a graduation subject.

\subsection*{Acknowledgements}
There are some people I would like to thank that contributed to my project
in various ways. This thesis would not have been possible without them.

\paragraph{}
First, I would like to thank Brett Kelly, who was CTO at VNU Vacature Media at
the time I started the Master's programme, and Rutger Pannekoek, who is CTO of
VNU Vacature Media since 2013. Without them, I wouldn't be able to follow the
Master's programme together with my job in the first place.

Second, I would like to thank my tutor at the University of Amsterdam, Magiel
Bruntink, for the help and support during this project.

I would also like to thank my parents, Johan and Carla Endenburg for always
supporting me in whatever I do. They always told me that ``when there's a will,
there's a way''.

Thanks go out to my colleagues
Bas,
Boudewijn,
Danny,
Eugene,
Ewout,
Greg,
Hugo,
Jan-Willem,
Jeroen,
Jerry,
Joeri,
Johan,
Marc,
Marvin,
Merlijn,
Michiel,
Nienke,
Kalle,
Leonard,
Silvester,
Tim,
and Wouter
(if I forget anyone, sorry for that), you were all great in supporting me in
conducting my research on the days that I was physically present, but mentally
absent.

Additionally, I would like to thank Siim Karus for helping me replicating his
research. He has been very helpful, especially in the beginning of the project,
in providing me his scripts and insights in his data.

Last, but definitely not least, big thanks to Christine Fr\"{o}ger, for all the
love, support, patience, and encouragements to complete this thesis and for
telling me to take a break every now and then.\\[2em]

\begin{flushright}
Martijn Endenburg\\
Amsterdam, The Netherlands\\
July, 2014
\end{flushright}
